---
title: "General Social Survey"
author: "Mairani"
date: "Summer 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: CAPPUN is independent of OWNGUN  
$H_A$: CAPPUN is not independent of OWNGUN  

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

CAPPUN is a categorical variable,   
OWNGUN is a categorical variable, so I use the cat-cat helper file.  

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
barchartGC(~CAPPUN + OWNGUN,data=GSSdata)
barchartGC(~CAPPUN + OWNGUN,data=GSSdata, type="percent")

```

The Plot barplots show that there is a higher frequency of people who favor the death penalty compared to those who oppose it.It also shows that if they oppose the penalty, they are more likely to not be a gun owner.    

#### Numerical Descriptive Results  

```{r}
table1 <- xtabs(~CAPPUN + OWNGUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```
This table shows that the largest group opposing CAPPUN are people that dont own a gun, they are 70.97 of the total group of opposers.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the Chi-squared test was p=0.02022.
This means we reject the null hypothesis,and conclude that opinion on capital punishment is dependent of gun ownership.  

The p-value for the fisher exact test was p= 0.01651, and the odds ratio was 2.271. This means that people who do not own a gun are 2.271 times more likely to oppose capital punishment. 

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

$H_0$: there is no difference in years of education(EDUC) between those who did or didn't vote in 2012(VOTE12).  

$H_A$: There is a difference in years of education(EDUC) between those who did or didn't vote in 2012(VOTE12).  

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results
Years of education (EDUC) is a numerical variable. It is a quantitative measure.  

Voting in 2012(VOTE12) is categorical,yes or no.    

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
It is numerical-categorical so we use the numerical-Cat helper file.  

```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

The boxplot shows that voting in 2012 is dependent of education.   

VOTE12 was higher as the education was higher. Those that didn't vote generally had lower years of education.  

#### Numerical Descriptive Results

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
  The numerical summary shows that for individuals who voted in 2012 (voted12), the lowest level of education among the group is 9 years.For those that didn't vote, it was 7 years.   
  Therefore, those that voted had more education.  
  
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

