---
title: "General Social Survey"
author: "Mairani"
date: "Summer 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---

```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```

# Question 1


The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.\
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: CAPPUN is independent of OWNGUN  

$H_A$: CAPPUN is not independent of OWNGUN    

## Methods


CAPPUN is a categorical variable,
OWNGUN is a categorical variable, so I use the cat-cat helper file.

## Results

### Descriptive Results

#### Graphical Descriptive Results


```{r}
barchartGC(~CAPPUN + OWNGUN,data=GSSdata)
barchartGC(~CAPPUN + OWNGUN,data=GSSdata, type="percent")

```

The Plot barplots show that there is a higher frequency of people who favor the death penalty compared to those who oppose it.It also shows that if they oppose the penalty, they are more likely to not be a gun owner.

#### Numerical Descriptive Results

```{r}
table1 <- xtabs(~CAPPUN + OWNGUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

This table shows that the largest group opposing CAPPUN are people that dont own a gun, they are 70.97 of the total group of opposers.


### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the Chi-squared test was p=0.02022. This means we reject the null hypothesis,and conclude that opinion on capital punishment is dependent of gun ownership.

The p-value for the fisher exact test was p= 0.01651, and the odds ratio was 2.271. This means that people who do not own a gun are 2.271 times more likely to oppose capital punishment.


# Question 2

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$: there is no difference in years of education(EDUC) between those who did or didn't vote in 2012(VOTE12).   


$H_A$: There is a difference in years of education(EDUC) between those who did or didn't vote in 2012(VOTE12).     

## Methods
Years of education (EDUC) is a numerical variable. It is a quantitative measure.

Voting in 2012(VOTE12) is categorical,yes or no.  

## Results

### Descriptive Results


#### Graphical Descriptive Results


It is numerical-categorical so we use the numerical-Cat helper file.

```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

The boxplot shows that voting in 2012 is dependent of education.

VOTE12 was higher as the education was higher. Those that didn't vote generally had lower years of education.

#### Numerical Descriptive Results

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```


The numerical summary shows that for individuals who voted in 2012 (voted12), the lowest level of education among the group is 9 years.For those that didn't vote, it was 7 years.  
Therefore, those that voted had more education.

### Inferential Results

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```


The Welch Two Sample t-test shows that the p-value = 1.248e-07 (0.0000001248). It is much smaller than 0.05 so we reject the null hypothesis.

# Question 3


Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

$H_0$: There is no difference between a persons performance on the wordsum test and their level of schooling.    


$H_A$: There is a difference between a persons performance on the wordsum test and their level of schooling.    

## Methods
  
WORDSUM is a numerical variable and DEGREE is a categorical variable, so I use the Numerical-Cat helper file.    


### Results

### Descriptive Results

#### Graphical Descriptive Results

```{r}
plot(WORDSUM~DEGREE,data=GSSdata)
```

The plot shows that the highest scorers (a score of 12) had a bachelors degree. They averaged a score around 10 and LT High School averaged around 6-8.  

#### Numerical Descriptive Results
```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

The numerical summary shows that LT High School had a lower minimum on the test with a score of 3.Graduates had a higher minimum of 7, indicating that a higher level of education results in a higher score on the WORDSUM test.

### Inferential Results
```{r}
anova=aov(WORDSUM~DEGREE,data=GSSdata)
summary(anova)
```

This shows that we reject the null hypothesis because the p-value is 0.0000002 which is smaller than 0.05.

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(7, 8.7, 2, 2))
plot(TukeyHSD(anova),las=2)
```
  
There is no difference between education in highschool and less than highschool on the test. However,there is a difference between jr. college and highschool. Overall, confidence levels seem to be impacted by education level. 

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion
For Question 1, the barchart shows that opposers of the penalty are more likely to not own a gun. The p-value for the Chi-squared test was p=0.02022,it was small, also proving that the null hypothesis is rejected because opinion on capital punishment is dependent of gun ownership.CAPPUN is not independent of OWNGUN.

For Question 2, the null-hypothesis,"There is no difference in years of education(EDUC) between those who did or didn't vote in 2012(VOTE12)" is rejected. The plot showed that VOTE12 was higher as the education was higher. Favstats also represented the years of education voters had. Voters had more years of education vs non-voters. Lastly, the t-test shows that the p-value = 0.0000001248,which is so small that it tells us we reject the null hypothesis.

For Question 3, the p-value was 0.0000002, so the null hypothesis was rejected. The plot and favstats showed that higher levels of education typically ranked higher on the WORDSUM test than highschoolers. Therefore, it was not proven that there is no difference between a persons performance on the wordsum test and their level of schooling.

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
